Ae. albopictus CPP manuscript

### 2019-11-06 MS.Rmd
added supp fig

### Mosquito population sampling and pre-experiment colony maintenance
Mosquito populations were sampled as larvae and pupae in the eastern USA and Japan in 2018, at latitudes ranging from `r min(pops$latitude)` to `r max(pops$latitude)`. Sampling locations are described in Table 1. Locations in the US were identical to locations sampled in 2008, except: . Japanese population samples were sent to Georgetown as F1 eggs (except for TAK, for which F2 eggs were provided). At least 50 male and 50 female mosquitoes were used to initiate the laboratory populations at Georgetown, except for SEN, which had 37 females and 39 males.

@@

Populations were maintained in the laboratory for 3 or 4 generations prior to the experiment at 21C and 80% relative humidity under a 16:8h light:dark cycle, following previous studies [@armbrusterGeographicVariationLarval2006; @urbanskiRapidAdaptiveEvolution2012]. Larvae were reared in distilled water and fed a slurry of dog food and brine shrimp. Adults were provided organic raisins as a carbohydrate source and blood meals from a human host. Initial laboratory generations of larvae were reared in plastic storage containers containing ~2.5 L of distilled water, with one population per container at initial densities of ~200-300. Pupae were then transferred into one adult cage per population. Populations generally had at least 50 female and 50 male adults per generation, but the JAC population experienced visibly high mortality during the F2 generation. The larval generation immediately prior to the experiment was reared in 15 cm plastic petri dishes containing 30-40 larvae each, 10 dishes per population, to match the larval rearing conditions of the experimental generation, which followed the methods of Urbanski et al 2012. Additional details of mosquito husbandry can be found in Supplementary File 1.

### Critical photoperiod determination
An overview of the procedure used to determine the critical photoperiod is as follows: Starting from the pupal stage, mosquitoes were divided among 12 different photoperiods (listed in Supplementary File 1) and allowed to produce eggs. The proportion of diapause eggs at each photoperiod was determined, and a dose-response curve was constructed with respect to hours of light and proportion of eggs in diapause. The critical photoperiod was calculated as the number of hours of light at which 50% of the eggs produced were in diapause, excluding the fraction of constitutively non-diapausing eggs (i.e. the fraction that did not enter diapause under unambigous short-day conditions, 8 hours of light). The experimental design and most of the equipment and experimental details were the same as those used in Urbanski et al 2012; additional details on all steps can be found in Supplementary File 1.

@@

Pupae from each population were collected, pooled across dishes, and split amongst thirteen 1.5 L adult cages, corresponding to 12 photoperiod cabinets and an additional short-day (8h light) incubator to provide additional data on unambiguous short-day diapause incidence. The total number of pupae added to each cage ranged from 52 to 106; pupae were not sexed and subsequent pupal and adult mortality was not monitored. Approximately one week after the last pupae were added, cages were blood-fed, and a dark oviposition cup half-filled with water and lined with unbleached paper towel was added to each cage. Oviposition papers were collected 5-7 days after blood-feeding and replaced every MWF; a total of four oviposition papers were collected from each cage. Several cages were fed again and an additional oviposition paper was collected; see Supplementary File 1 for details. Oviposition papers were dried gently and stored in the photoperiod cabinets from which they originated for 11 to 21 days until diapause determination.

@@

```{r}
plot(blockAB$yield ~ blockAB$TotalLarvaeEmbryos, xlab="Number of viable eggs on oviposition paper", ylab="Viable eggs/initial egg count on paper")
```

>>>>

### Mosquito population sampling and pre-experiment colony maintenance
Mosquito populations were sampled as larvae and pupae in the eastern USA and Japan in 2018, at latitudes ranging from `r min(pops$latitude)` to `r max(pops$latitude)`. Sampling locations are described in Table 1. Locations in the US were identical to locations sampled in 2008. Japanese population samples were sent to Georgetown as F1 eggs (except for TAK, for which F2 eggs were provided). At least 50 male and 50 female adult mosquitoes were used to initiate all populations in the laboratory at Georgetown, except for SEN, which had 37 females and 39 males.

@@

Populations were maintained in the laboratory for 3 or 4 generations prior to the experiment at 21C and 80% relative humidity under a 16:8h light:dark cycle, following previous studies [@armbrusterGeographicVariationLarval2006; @urbanskiRapidAdaptiveEvolution2012]. Larvae were reared in distilled water and fed a slurry of dog food and brine shrimp. Adults were provided organic raisins as a carbohydrate source and blood meals from a human host. Initial laboratory generations of larvae were reared in plastic storage containers containing ~2.5 L of distilled water, with one population per container at initial densities of ~200-300. Pupae were then transferred into one adult cage per population. Populations generally had at least 50 female and 50 male adults per generation, but the JAC population experienced visibly high mortality during the F2 generation. The larval generation immediately prior to the experiment was reared in 15 cm plastic petri dishes containing 30-40 larvae each, 10 dishes per population, to match the larval rearing conditions of the experimental generation, which followed the methods of Urbanski et al 2012. Additional details of mosquito husbandry prior to the experiment can be found in Supplementary File 1.

### Critical photoperiod determination
An overview of the procedure used to determine the critical photoperiod is as follows: Starting from the pupal stage, mosquitoes were divided among 12 different photoperiods (listed in Supplementary File 1) and allowed to produce eggs. The proportion of diapause eggs at each photoperiod was determined, and a dose-response curve was constructed with respect to hours of light and proportion of eggs in diapause. The critical photoperiod was calculated as the number of hours of light at which 50% of the eggs produced were in diapause, excluding the fraction of constitutively non-diapausing eggs (i.e. the fraction that did not enter diapause under unambigous short-day conditions, 8 hours of light). This experimental design and most of the equipment and experimental details were the same as those used in Urbanski et al 2012; additional details on all steps can be found in Supplementary File 1.

@@

Pupae from each population were collected, pooled across dishes, and split amongst thirteen 1.5 L adult cages, corresponding to twelve photoperiod cabinets and an additional short-day (8h light) incubator to provide additional data on unambiguous short-day diapause incidence. The total number of pupae added to each cage ranged from 52 to 106; pupae were not sexed and subsequent pupal and adult mortality was not monitored. Approximately one week after the last pupae were added, cages were blood-fed, and a dark oviposition cup half-filled with water and lined with unbleached paper towel was added to each cage. Oviposition papers were collected 5-7 days after blood-feeding and replaced every MWF; a total of four oviposition papers were collected from each cage. Several cages were fed again and an additional oviposition paper was collected; see Supplementary File 1 for details. Oviposition papers were dried gently and stored in the photoperiod cabinets from which they originated for 11 to 21 days until diapause determination.

@@

```{r}
plot(blockAB$yield ~ blockAB$TotalLarvaeEmbryos, xlab="Number of viable eggs on oviposition paper", ylab="Viable eggs/initial egg count on paper")
```
```{r}
#comparing CPP values calculated with and without filtering out unreliable oviposition papers
cpp_filtered<-read.csv("cpp_calcs_filtered.csv") %>% filter(year==2018)
cpp_unfiltered<-read.csv("cpp_calcs_unfiltered.csv") %>% filter(year==2018)
names(cpp_filtered)[5]<-"CPP_est_filt"
names(cpp_unfiltered)[5]<-"CPP_est_unfilt"
cpp_compare<-left_join(cpp_filtered,cpp_unfiltered,by="pop")
ggplot(cpp_compare, aes(x=CPP_est_unfilt, y=CPP_est_filt)) + geom_point() + geom_abline(slope=1, intercept=0)
